---
title: 'Seminar 9 (Week 10), Audio and Image Data'
author: "Melissa Sands"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting Up

## Functions

First, we load some re-usable functions for our activity. You can inspect, edit, and add to the functions by opening the relevant .R script. But for now, we can just source them.

```{r functions}
source("R//functions//functions.R")
```

Let's now load the necessary libraries for this activity:

```{r libraries}
# start by checking for and installing packages:
package_check(c("devtools", "ggplot2", "dplyr", "stringr", "tidylog", "rvest", "tuneR", "seewave", "imager"))

# devtools for installing from github
# ggplot2 for visualisation
# dplyr for data manipulation
# stringr for string manipulation
# tidylog to keep track of our data manipulation
# rvest for web scraping
# tuneR for audio loading, analysis, etc.
# seewave for audio visualisation
# imager for image loading, analysis, etc. Also works for video, but we won't do that today.
# magick -- we won't call this but it is an alternative suite to imager, similar types of tools, and interchanageable/complementary

```

# Audio Data

## Downloading and Ingesting Data

```{r audio_download}
# set url of audio file:
url <- "https://s3.amazonaws.com/RE-Warehouse/w/winston_churchill_speeches_and_radio_broadcasts_1940-06-04_we_shall_never_surrender.mp3"

# extract the file name from the url:
file_name <- basename(url)
img_grabber(file_name, url) # check the functions.R file to see what this function does

```

Now let's ingest the audio file:

```{r audio_ingest}
# ingest the audio file:
churchill <- readMP3(paste0("data/", file_name))

# look at it
churchill

```

Now let's clip the audio to a smaller segment, so we can analyse it:

```{r audio_clip}
# clip the audio file to a 30 second segment
churchill_clip <- extractWave(churchill, from=670, to=700, xunit='time')

# play the clip -- you may need to authorize/download something here, depending on your system. 
tuneR::play(churchill_clip)
     
```

And now, conduct some visualisation:

```{r audio_viz}
# visualise amplitudes or waveform (loudness)
tibble(seconds = (1:length(churchill_clip@left))/churchill_clip@samp.rate,
             amplitude = churchill_clip@left) %>%
  ggplot(aes(x = seconds, y = amplitude)) +   
    geom_line() +
    theme_minimal()

# create a spectrogram (frequency and amplitude)
seewave::spectro(churchill_clip)

# closer inspection:
churchill_clip_1s <- extractWave(churchill_clip, from=1, to=2, xunit='time')
seewave::spectro(churchill_clip_1s)

```

# Image Data

Let's turn to image data. We will start by by just exploring image data a little. 

First, let's grab some images from the web. This can take a while if we have many images, so start with a small set:

```{r image_ingestion}
# A couple of political images from Wikimedia Commons.
# These are stable, openly licensed direct image URLs.
urls <- c(
  "https://upload.wikimedia.org/wikipedia/commons/4/49/Occupy_Wall_Street_Crowd_Size_2011_Shankbone.JPG",   # Occupy Wall Street crowd, NYC 2011
  "https://upload.wikimedia.org/wikipedia/commons/3/3d/View_of_Crowd_at_1963_March_on_Washington.jpg"     # March on Washington crowd, 1963
)

# load the images into R as a list:
img_lst <- list()

for(i in seq_along(urls)){
  img_lst[[i]] <- try(load.image(urls[i]))
}

# remove any that failed to load:
img_lst <- img_lst[!sapply(img_lst, inherits, "try-error")]

# how many loaded successfully?
length(img_lst)
```

Let's take a look at our images. 

```{r image_inspection}
# plot all images to get an overview:
for(i in seq_along(img_lst)) plot(img_lst[[i]], main = paste("Image", i), axes = FALSE)
```

```{r grayscale_conversion}
# image 1 is in colour; convert to grayscale:
img1_grayscale <- grayscale(img_lst[[1]])
plot(img1_grayscale)

```

```{r image_analysis}
# conduct edge detection (how pixel intensity changes from pixel to pixel within a small area)
# edge detection relies on image gradients, which imager returns via:
gr1 <- imgradient(img_lst[[2]],"xy")
gr1
plot(gr1, layout="row")

```

# Self-Directed Task:

Find either an interesting audio clip or an interesting image (or set of images) online. Download the files, ingest them, and do some exploratory analyses. 
